Towards a miniKanren with fair search strategies

### 2019-05-23 Paper/paper.tex
raise one reverso
have shown -> have demonstrated
eliminate abbrev. in Abstract; repeato -> repeat^o

\newcommand{\veryrecursiveo}{\texttt{very-recursive$^o$}}
\newcommand{\appendo}{\texttt{append$^o$}}
\newcommand{\reverso}{\texttt{revers$^o$}}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can
be fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search (\DFSi),
the balanced interleaving depth-first search (\DFSbi), the fair depth-first
search (\DFSf), and the standard breadth-first search (\BFSser).
\DFSbi{} and \DFSf{} are new. And we present a new, more efficient
and simpler implementation of \BFSser. Using quantitative evaluation, 
we argue that
\DFSbi{} and \DFSf{} are competitive alternatives to \DFSi, 
and that our \BFSser{} implementation is more efficient than the current one.

@@

A subtlety arises 
when a \conde{} contains many clauses: not every clause has an 
equal chance of contributing to the result. As an example, consider the following 
relation \texttt{repeato} and its invocation.

@@

Next, consider the following disjunction of invoking \texttt{repeato} with four 
different letters.

@@

Now we elaborate fairness by running more queries about \texttt{repeato}.

@@

Under the hood, the \conde{} and the call to \texttt{repeato} are connected by 
\conj{}. The \conde{} goal outputs a search space including four trivially 
different states. 
Applying the next conjunctive goal, \texttt{(repeato x q)}, produces four 
trivially different search spaces.
In the examples above, all search strategies allocate more computational 
resources to the search space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each search space. For example,

@@

\citet{kiselyov2005backtracking} have shown that a \emph{MonadPlus} hides in 
implementations of logic programming system. Our \BFSser{} implementation is 
not an exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, 
but takes arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, 
and \texttt{mplus}, respectively.

@@

Here we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means ``running out of 
memory.'' The first two benchmarks are from 
TRS2. \texttt{reverso} is from 
\citet{rozplokhas2018improving}. The next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevents closures from being inside a list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is from \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that \DFSi{} 
would 
perform badly in the second case because elimination rules complicate the 
problem when synthesizing (i.e., our evaluation supports our conjecture.)

@@

\NOTE[DPF]{appendo is not raised, though I tried.}

>>>>

\newcommand{\veryrecursiveo}{\texttt{very-recursive$^o$}}
\newcommand{\appendo}{\texttt{append$^o$}}
\newcommand{\reverso}{\texttt{revers$^o$}}
\newcommand{\repeato}{\texttt{repeat$^o$}}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can
be fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.
The two non-standard depth-first searches are new. And we present a new, more 
efficient and simpler implementation of the standard breadth-first search. Using 
quantitative evaluation, we argue that the two new depth-first searches are 
competitive alternatives to the standard one, 
and that our standard breadth-first search implementation is more efficient than 
the current one.

% We describe fairness levels in disjunction and conjunction
% implementations.  Specifically, a disjunction implementation can be
% fair, almost-fair, or unfair. And a conjunction implementation can
% be fair or unfair.  We compare the fairness level of four search
% strategies: the standard miniKanren interleaving depth-first search (\DFSi),
% the balanced interleaving depth-first search (\DFSbi), the fair depth-first
% search (\DFSf), and the standard breadth-first search (\BFSser).
% \DFSbi{} and \DFSf{} are new. And we present a new, more efficient
% and simpler implementation of \BFSser. Using quantitative evaluation, 
% we argue that
% \DFSbi{} and \DFSf{} are competitive alternatives to \DFSi, 
% and that our \BFSser{} implementation is more efficient than the current one.

@@

A subtlety arises 
when a \conde{} contains many clauses: not every clause has an 
equal chance of contributing to the result. As an example, consider the following 
relation \repeato{} and its invocation.

@@

Next, consider the following disjunction of invoking \repeato{} with four 
different letters.

@@

Now we elaborate fairness by running more queries about \repeato{}.

@@

Under the hood, the \conde{} and the call to \repeato{} are connected by 
\conj{}. The \conde{} goal outputs a search space including four trivially 
different states. 
Applying the next conjunctive goal, \texttt{(repeato x q)}, produces four 
trivially different search spaces.
In the examples above, all search strategies allocate more computational 
resources to the search space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each search space. For example,

@@

\citet{kiselyov2005backtracking} have demonstrated that a \emph{MonadPlus} 
hides in 
implementations of logic programming system. Our \BFSser{} implementation is 
not an exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, 
but takes arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, 
and \texttt{mplus}, respectively.

@@

Here we compare the efficiency of search strategies. A concise 
description is in Table~\ref{compare-efficiency}. A hyphen means ``running out of 
memory.'' The first two benchmarks are from 
TRS2. \reverso is from 
\citet{rozplokhas2018improving}. The next two benchmarks 
about quine are modified from a similar test case in \citet{byrd2017unified}. 
The modifications are made 
to circumvent the need for symbolic constraints (e.g. $\neq$, 
\texttt{absent$^o$}). Our version generates de 
Bruijnized expressions and prevents closures from being inside a list. The two 
benchmarks differ in the \conde{} clause order of their relational 
interpreters. 
The last two 
benchmarks are about synthesizing expressions that evaluate to \texttt{'(I love 
you)}. This benchmark is from \citet{byrd2017unified}. Again, the 
sibling benchmarks differ in the \conde{} clause order of their relational 
interpreters. The first one 
has elimination rules (i.e. application, \texttt{car}, and \texttt{cdr}) at the 
end, while the other has them at the beginning. We conjecture that \DFSi{} 
would 
perform badly in the second case because elimination rules complicate the 
problem when synthesizing (i.e., our evaluation supports our conjecture.)

@@

\NOTE[DPF]{appendo is not raised, though I tried.}
\NOTE[LKC]{Which appendo is not raised? I couldn't find one.}
